Switch equivalence and weak prime domains for left-linear DPO rewriting systems

### 2025-04-17 full_version/esempi.tex
aggiunto capitolo

% !TeX root = ./fusione.tex

\chapter{Independence in rewriting}

In this chapter we discuss the notion of independence between two
consecutive rewriting steps, a fundamental ingredient of the rewriting
theory, which comes into play when viewing a sequence of steps as a
concurrent computation. We observe that moving from linear to
left-linear rules leads to the failure of various basic properties of
the classical notion of independence used in the linear setting and we
single out a framework in which these can be re-established, possibly
in a weakened form.

\section{Sequentially independent and switchable derivations }\label{subsec:switch}

Sequential independence is the canonical notion of independence in the DPO approach.

\begin{definition}[Sequential independence]
	\label{de:sequential-independence}
	Let $\X$ be an $\mathcal{M}$-adhesive category 
	and $(\X, \R)$ a left-linear rewriting system. 
	Let also
	$\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation of length $2$, with
	$\dder{D}_i$ using rule $\rho_i = (l_i,r_i)$.  We say that $\dder{D}_0$ and
	$\dder{D}_1$ are
	\emph{sequentially independent}
	if there is a pair of arrows
	$i_0\colon R_0\to D_1$ and $i_1\colon L_1\to D_0$ such that the
	following diagram commutes. In this case $(i_0,i_1)$ is called an  \emph{independence
		pair}.
	\[\xymatrix@R=20pt@C=20pt{L_0 \ar[d]_{m_0}&& K_0
		\ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
		\ar@{.>}@/^.25cm/[drrr]|(.32)\hole_(.4){i_0}
		\ar[dr]|(.33)\hole_{h_0} && L_1 \ar@{.>}@/_.25cm/[dlll]^(.4){i_1}
		\ar[dl]|(.33)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1}
		\ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0}
		D_0 \ar[rr]_{g_0}&& G_1 && \ar@{>->}[ll]^{f_1} D_1
		\ar[rr]_{g_1}&& G_2}
	\]
\end{definition}

\begin{remark}\label{rem:seqabs}
Let $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ be an abstraction equivalence between $\der{D}=\{\dder{D}_i\}_{i=0}^1$ and $\der{D}'=\{\dder{D}'_i\}_{i=0}^1$. If there is an independence pair between $(i_0, i_1)$ between $\dder{D}_0$ and $\dder{D}_1$, then $(\phi_{D_1}\circ i_0, \phi_{D_0}\circ i_1)$ is an independence pair between $\dder{D}'_0$ and $\dder{D}'_1$. Indeed we have
\[\begin{split}
g'_0\circ \phi_{D_0}\circ i_1&=\phi_{G_1}\circ g_0\circ i_1\\&=\phi_{G_1}\circ m_1\\&= m'_1
\end{split}\qquad \begin{split}
f'_1\circ \phi_{D_1}\circ i_0&=\phi_{G_1}\circ f_1\circ i_0\\&=\phi_{G_1}\circ h_0\\&= h'_1
\end{split}\]
\end{remark}

\begin{remark}\label{rem:uni}
	It is worth mentioning that if $(\X, \R)$ is a linear rewriting system then two derivations $\dder{D}_0$ and $\dder{D}_1$ can have at most one independence pair. Indeed if $(i_0,i_1)$ and $(i'_0, i'_1)$ are independence pairs, then
	\[g_0\circ i_1 = g_0\circ i'_1 \qquad f_1\circ i_0=f_1\circ i'_0\]
	But in linear systems $g_0$ and $f_1$ are both monos, entailing $i_0=i'_0$ and $i_1=i'_1$.
\end{remark}

Intuitively, the existence of the independence pair
captures the possibility of switching the application of the two rules:
$f_0 \circ i_1$ is a match for $\rho_1$ in $G_0$ and the
existence of $i_0 : R_0 \to D_1$ means that $\rho_1$ does not delete
items in the image of $K_0$, thus $\rho_0$ can be applied after
$\rho_1$. The following definition formalise what it means for a derivation to be the switch of another.

\begin{definition}[Pre-switch] \label{def:switch}
	Let $\X$ be an $\mathcal{M}$-adhesive category 
	and $(\X, \R)$ a left-linear rewriting system. 
	Let also
	$\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation made
	of two sequentially independent derivations $\dder{D}_0$,
	$\dder{D}_1$ with independence pair $(i_0, i_1)$, as in the following diagram. 
		\[\xymatrix@R=20pt@C=20pt{L_0 \ar[d]_{m_0}&& K_0
		\ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
		\ar@{>}@/^.25cm/[drrr]|(.32)\hole_(.4){i_0}
		\ar[dr]|(.33)\hole_{h_0} && L_1 \ar@{>}@/_.25cm/[dlll]^(.4){i_1}
		\ar[dl]|(.33)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1}
		\ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0}
		D_0 \ar[rr]_{g_0}&& G_1 && \ar@{>->}[ll]^{f_1} D_1
		\ar[rr]_{g_1}&& G_2}
	\]

	A \emph{pre-switch} of $\der{D}$ along $(i_0,i_1)$ is
	a derivation $\der{E}=\{\dder{E}_i\}_{i=0}^1$, between the same objects and using the same rules
	in reverse order, as in the diagram on the right below
	\[
	\xymatrix@R=20pt@C=20pt{
		{L_1} \ar[d]_{m_0'}
		&&  {K_1} \ar@{>->}[ll]_{l_1} \ar[r]^{r_1} \ar[d]_{k_0'}
		&  {R_1} \ar[dr]|(.33)\hole_{h_0'}  \ar@/^.25cm/@{.>}_(.4){i_0'}|(.32)\hole[drrr]
		& & 
		{L_0}\ar[dl]|(.33)\hole^{m_1'} \ar@/_.25cm/@{.>}^(.4){i_1'}[dlll] 
		&  {K_0} \ar@{>->}[l]_{l_0} \ar[rr]^{r_0} \ar[d]^{k_1'}
		& & {R_0} \ar[d]^{h_1'} \\		
		{G_0}
		& & {D_0'} \ar@{>->}[ll]^{f_0'} \ar[rr]_{g_0'}
		& &  {G'_1} 
		& &  {D_1'} \ar@{>->}[ll]^{f_1'} \ar[rr]_{g_1'}
		& & {G_2}  }
	\]
	such that there is an independence pair $(i_0', i_1')$ between
	$\dder{E}_0$ and $\dder{E}_1$ and 
\[
	m_0=f_0' \circ i_1' \qquad 
	h_1=g_1' \circ i_0' 
		\qquad m_0'= f_0 \circ i_1
		\qquad h_1'= g_{1}\circ i_0\]

\end{definition}

\begin{example}	\label{ex:seq-ind}
	Consider the rewriting system  $(\cat{Graph}, \R)$ and the derivation $\der{D}$ introduced in \Cref{ex:1}.	Steps $\dder{D}_1$ and $\dder{D}_2$ are clearly sequential
	independent via the arrows $R_1 \to D_2$, $L_2 \to D_1$
	mapping nodes according to their numbering.

	A pre-switch of $\der{D}$ along such independence pair is the derivation
	$\der{E}$ depicted below. Instead,  the first two
	steps $\dder{D}_0$ and $\dder{D}_1$ are not sequential independent. Intuitively, this happens because $\rho_1$ uses the node generated by $\rho_0$ to produce a self-loop.
\begin{center}
	\begin{tikzpicture}[node distance=2mm, font=\small,  baseline=(current bounding box.center)]      
		\node (L1) at (0,2) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {};
				\node at (0,0) [node, label=below:$1$] (1) {} ;
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		\node [right=of L1] (K1) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$1$] (1) {};
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		%     \node [above=of K1] {$\rho_0$};
		\node [above] at (K1.north) {$\rho_0$};
		\node [right=of K1](R1) {
			\begin{tikzpicture}
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$1$] (1) {};
				\node at (.5,0) [node, label=below:$2$] (2) {}; 
				\draw[coloredge] (1) to[out=20, in=160] (2);
				% 
				\pgfBox
			\end{tikzpicture}
		};
		\path (K1) edge[->] node[trans, above] {} (L1);
		\path (K1) edge[->] node[trans, above] {} (R1);

		\node at (4,2) (L3) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$1$] (1) {} ;
				\node at (0.5,0) [node, label=below:$2$] (2) {} ;
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		\node [right=of L3] (K3) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$1$] (1) {} ;
				\node at (0.5,0) [node, label=below:$2$] (2) {} ;
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		%     \node [above=of K3] {$\rho_2$};
		\node [above] at (K3.north) {$\rho_2$};
		\node [right=of K3] (R3) {
			\begin{tikzpicture}
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$12$] (12) {};
				% 
				\pgfBox
			\end{tikzpicture}
		};
		\path (K3) edge[->] node[trans, above] {} (L3);
		\path (K3) edge[->] node[trans, above] {} (R3);

		\node at (8,2) (L2) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$2$] (1) {} ;
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		\node [right=of L2] (K2) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$2$] (1) {};
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		%     \node [above=of K2] {$\rho_1$};
		\node [above] at (K2.north) {$\rho_1$};

		\node [right=of K2] (R2) {
			\begin{tikzpicture}
				\node at (0,0) [node, label=below:$2$] (1) {}
				edge [in=55, out=85, loop] ();
				% 
				\pgfBox
			\end{tikzpicture}
		};
		\path (K2) edge[->] node[trans, above] {} (L2);
		\path (K2) edge[->] node[trans, above] {} (R2);

		%%%%%% second row
		\node at (0,0) (G1) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {};
				\node at (0,0) [node, label=below:$1$] (1) {} ;
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		\node [right=of G1] (D1) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$1$] (1) {};
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		\node at (3,0) (G2) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$1$] (1) {} ;
				\node at (0.5,0) [node, label=below:$2$] (2) {} ;
				\draw[coloredge] (1) to[out=20, in=160] (2);
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		\path (D1) edge[->] node[trans, above] {} (G1);
		\path (D1) edge[->] node[trans, above] {} (G2);
		\path (L1) edge[->] node[trans, above] {} (G1);
		\path (K1) edge[->] node[trans, above] {} (D1);
		\path (R1) edge[->] node[trans, above] {} (G2);

		\node at (5.5,0) (D2) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {};
				\node at (0,0) [node, label=below:$1$] (1) {} ;
				\node at (0.5,0) [node, label=below:$2$] (2) {} ;
				\draw[coloredge] (1) to[out=20, in=160] (2);
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		\node at (7.4,0) (G3) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$12$] (12) {}
				edge [in=125, out=155, colorloop] ();
				% 
				\pgfBox
			\end{tikzpicture} 
		};

		\path (D2) edge[->] node[trans, above] {} (G2);
		\path (D2) edge[->] node[trans, above] {} (G3);
		\path (L3) edge[->] node[trans, above] {} (G2);
		\path (K3) edge[->] node[trans, above] {} (D2);
		\path (R3) edge[->] node[trans, above] {} (G3);

		\node at (9.1,0) (D3) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$12$] (12) {}
				edge [in=125, out=155, colorloop] ();
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		\node [right=of D3] (G4) {
			\begin{tikzpicture}
				% 
				\node at (0,0) [node, label=below:$12$] (12) {}
				edge [in=55, out=85, loop] ()
				edge [in=125, out=155, colorloop] ();
				% 
				\pgfBox
			\end{tikzpicture} 
		};

		\node[font=\scriptsize, below] at (G1.south) {$G_0$};
		\node[font=\scriptsize, below] at (D1.south) {$D_0$};      
		\node[font=\scriptsize, below] at (G2.south) {$G_1$};
		\node[font=\scriptsize, below] at (D2.south) {$D_1'$};      
		\node[font=\scriptsize, below] at (G3.south) {$G_2'$};
		\node[font=\scriptsize, below] at (D3.south) {$D_2'$};      
		\node[font=\scriptsize, below] at (G4.south) {$G_2$};

		\path (D3) edge[->] node[trans, above] {} (G3);
		\path (D3) edge[->] node[trans, above] {} (G4);
		\path (L2) edge[->] node[trans, above] {} (G3);
		\path (K2) edge[->] node[trans, above] {} (D3);
		\path (R2) edge[->] node[trans, above] {} (G4);
	\end{tikzpicture}
	%
\end{center}
	\end{example}

\begin{remark}\label{rem:switch}

Let $\der{E}=\{\dder{E}_i\}_{i=0}^1$ be a pre-switch for $\dder{D}_0$
and $\dder{D}_1$ along an independence pair $(i_0, i_1)$, then by definition we have an independence pair $(i'_0, i'_1)$ between $\dder{E}_0$ and $\dder{E}_1$ such that
\[
m_0=f_0' \circ i_1'
\quad h_1=g_1' \circ i_0'
\quad m_0'= f_0 \circ i_1
\quad h_1'= g_{1}\circ i_0\]

 The derivation $\der{D}$ uses the same rule of $\der{E}$ in reverse order, thus it is a switch for $\der{E}$ along $(i_0', i_1')$.   Moreover, the same argument shows that if $(\der{E}, \alpha, \omega)$ is a decorated switch for $(\der{D}, \alpha, \omega)$ along $(i_0, i_1)$, then the latter is a decorated switch for the former along $(i'_0, i'_1)$.
\end{remark}

We start by examining how pre-switches interact with abstraction equivalence. A first property that we can establish in that direction is uniqueness of pre-switches up to abstraction equivalence.

\begin{lemma}[Uniqueness of pre-switches]
	\label{thm:switch_uni}
	Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^1$ be a derivation and suppose
	that both $\der{E}=\{\dder{E}_i\}_{i=0}^1$ and
	$\der{F}=\{\dder{F}_i\}_{i=0}^1$ are pre-switches of $\dder{D}_0$ and
	$\dder{D}_1$ along the same independence pair $(i_0, i_1)$. Then there is a unique abstraction equivalence $\{\phi_X\}_{X\in \Deltamin(\der{E})}$ between $\der{E}$ and $\der{F}$, whose first component is $\id{G_0}$. Moreover, 
	\[\phi_{D_{\der{E}, 0}}\circ j_1=a_1 \qquad \phi_{D_{\der{E}, 1}}\circ j_0=a_0\]
\end{lemma}

\begin{proof}
	By definition of pre-switch $\der{E}_0$ and $\der{F}_0$ have the same match $f_0\circ i_1$ and that $\der{E}_1$ and $\der{F}_1$ have the same comatch $g_1\circ i_0$. Thus we get the solid part of the diagram below. If we are able to construct the dotted arrow and prove that they are isomorphism we are done.
	\[\xymatrix@C=22pt{G_{0} && \ar@{>->}[ll]_{f_{\der{F},0}} D_{\der{F},0} \ar[rr]^{g_{\der{F},0}}&& G_{\der{F},1} && \ar@{>->}[ll]_{f_{\der{F},1}} D_{\der{F},1} \ar[rr]^{g_{\der{F},1}}&& G_{2}\\L_1 \ar[d]^{f_0\circ i_1} \ar[u]_{f_0\circ i_1}&& K_{1} \ar[d]^{k_{\der{E},0}} \ar[u]_{k_{\der{F},0}}\ar@{>->}[ll]_{l_1} \ar[r]^{r_1} & R_{1} \ar@/^.35cm/[drrr]|(.315)\hole_(.4){j_0} \ar[dr]|(.3)\hole_{h_{\der{E},0}} \ar@/_.35cm/[urrr]|(.315)\hole^(.4){a_0} \ar[ur]|(.3)\hole^{h_{\der{F},0}}& &L_{0} \ar@/_.35cm/[dlll]^(.4){j_1} \ar[dl]|(.3)\hole^{m_{\der{E},1}} \ar@/^.35cm/[ulll]_(.4){a_1} \ar[ul]|(.3)\hole_{m_{\der{F},1}}& K_{0} \ar[d]_{k_{\der{E},1}} \ar[u]^{k_{\der{F},1}}\ar@{>->}[l]_{l_{0}} \ar[rr]^{r_0} && R_{0} \ar[d]_{g_1\circ i_0} \ar[u]^{g_1\circ i_0}\\ \ar@/^.5cm/[uu]^{\id{G_0}} G_{0} && \ar[ll]^{f_{\der{E},0}} D_{\der{E},0} \ar@{.>}@/^.5cm/[uu]^(.33){\phi_{0}}|\hole \ar[rr]_{g_{\der{E},0}}&& G_{\der{E},1} \ar@{.>}[uu]^(.76){\psi_1}|(.45)\hole |(.55)\hole&& \ar@{>->}[ll]^{f_{\der{E},1}} D_{\der{E},1} \ar@{.>}@/_.5cm/[uu]_(.33){\phi_{1}}|\hole\ar[rr]_{g_{\der{E},1}}&& G_{2}\ar@{.>}@/_.5cm/[uu]_{\psi_2}}\]

	Now, by \Cref{prop:unique} we know that there are two unique isomorphisms $\phi_0\colon D_{\der{E},0}\to D_{\der{F},0}$ and $\psi_1 \colon G_{\der{E},1}\to G_{\der{F},1}$ such that the diagram below commutes.
	\[\xymatrix{G_{0} & \ar@{>->}[l]_{f_{\der{F},0}} D_{\der{F},0} \ar[r]^{g_{\der{F},0}}& G_{\der{F},1} \\L_1 \ar[d]^{f_0\circ i_1} \ar[u]_{f_0\circ i_1}& K_{1} \ar[d]^{k_{\der{E},0}} \ar[u]_{k_{\der{F},0}}\ar[l]_{l_1} \ar[r]^{r_1} & R_{1}  \ar[d]_{h_{\der{E},0}}  \ar[u]^{h_{\der{F},0}}\\ \ar@/^.5cm/[uu]^{\id{G_0}} G_{0} & \ar@{>->}[l]^{f_{\der{E},0}} D_{\der{E},0} \ar@/^.5cm/[uu]^(.33){\phi_{0}}|\hole \ar[r]_{g_{\der{E},0}}& G_{\der{E},1} \ar@/_.5cm/[uu]_{\psi_1}}\]

	But then we have
	\begin{align*}
		f_{\der{F},0}\circ \phi_0\circ j_1&=f_{\der{E},0} \circ j_1\\&=m_0\\&=f_{\der{F},0}\circ a_1 \end{align*}
	which, since $f_{\der{D},0}$ is mono, entails $a_1= \phi_0\circ j_1$.	This equality, in turn, gives us that
	\begin{align*}
		\psi_1\circ m_{\der{E},1}&=\psi_1 \circ g_{\der{E},0}\circ j_1\\&=g_{\der{F},0}\circ \phi_0\circ j_1\\&=g_{\der{F},0}\circ a_1\\&=m_{\der{F},1}
	\end{align*}

	Since $\psi_1$ is an isomorphism, the square below is a pushout. 
	\[\xymatrix@C=35pt{K_0 \ar[d]_{k_{\der{E},1}} \ar@{>->}[r]^{l_0}& L_0 \ar[d]^{\psi_1\circ m_{\der{E},1}}\\D_{\der{E},1} \ar@{>->}[r]_{\psi_1\circ f_{\der{E},1}}& G_{\der{F}, 1}}\]

	Moreover, we have already shown that $\psi_1\circ m_{\der{E},1}=m_{\der{F},1}$. Thus, again by \Cref{prop:unique} there are the dotted isomorphisms $\phi_1\colon D_{\der{E},1}\to D_{\der{F},1}$ and $\psi_2\colon G_2\to G_2$ in the diagram below and such arrows are unique.
	\[\xymatrix{G_{\der{F},1} & \ar@{>->}[l]_{f_{\der{F},1}} D_{\der{F},1} \ar[r]^{g_{\der{F},1}}& G_2 \\L_0 \ar[d]^{m_{\der{E},1}} \ar[u]_{m_{\der{F},1}}& K_{0} \ar[d]^{k_{\der{E},1}} \ar[u]_{k_{\der{F},1}}\ar@{>->}[l]_{l_1} \ar[r]^{r_1} & R_{0}  \ar[d]_{g_1\circ i_0}  \ar[u]^{g_1\circ i_0}\\ \ar@/^.5cm/[uu]^{\psi_1} G_{\der{E},1} & \ar@{>->}[l]^{f_{\der{E},1}} D_{\der{E},1} \ar@{.>}@/^.5cm/[uu]^(.33){\phi_{1}}|\hole \ar[r]_{g_{\der{E},1}}& G_2 \ar@{.>}@/_.5cm/[uu]_{\psi_2}}\]

	To conclude, it is now enough to notice that
	\begin{align*}
		f_{\der{F},1} \circ \phi_1\circ j_0&=\psi_1\circ f_{\der{E},1}\circ j_0\\&=\psi_1\circ h_{\der{E,0}}\\&=h_{\der{F},0}\\&=f_{\der{F},1}\circ a_0
	\end{align*}
	allowing us to deduce $ \phi_1\circ j_0=a_0$.
\end{proof}

\begin{example}\label{ex:abs}
	Let $\X$ be the category of $\Set$ and, besides the terminal obect $1$, take the sets
	\[2=\{0,1\} \qquad 3=\{0,1,2\} \qquad 4=\{0,1,2,3\}\]

Let us consider the DPO-rewriting system given by the (linear) rules below.
		\[\xymatrix{1 & 1\ar@{>->}[l]_{\id{1}} \ar@{>->}[r]^{\id{1}} & 1 & 1 & 1\ar@{>->}[r]^{\delta^2_0} \ar@{>->}[l]_{\id{1}} & 2}\]

We then have the derivation $\der{D}$ below, composed by two sequentially independent steps.
		\[\xymatrix@R=20pt@C=20pt{1 \ar@{>->}[d]_{\delta^3_0}&& 1
	\ar@{>->}[d]_{\delta^3_0}\ar@{>->}[ll]_{\id{1}} \ar@{>->}[r]^{\id{1}} & 1
	\ar@{>->}@/^.25cm/[drrr]|(.32)\hole_(.4){\delta^3_0}
	\ar[dr]|(.33)\hole_{\delta^3_0} && 1 \ar@{>->}@/_.25cm/[dlll]^(.4){\delta^3_0}
	\ar@{>->}[dl]|(.33)\hole^{\delta^3_0}& 1 \ar@{>->}[d]^{\delta^3_0}\ar@{>->}[l]_{\id{1}}
	\ar@{>->}[rr]^{\delta^2_0} && 2 \ar@{>->}[d]^{i_2}\\3 && \ar@{>->}[ll]^{\id{3}}
	3 \ar@{>->}[rr]_{\id{3}}&& 3 && \ar@{>->}[ll]^{\id{3}} 3
	\ar@{>->}[rr]_{i_3}&& 4}
\]
where $i_3\colon 3\mto 4$ is the obvious inclusion, while $i_2\colon 2\mto 4$ sends $0$ to $0$ and $1$ to $3$.	 Now, we can construct two pre-switches for $\dder{D}_0$ and $\dder{D}_1$ along $(\delta_0, \delta_0)$. Indeed, on the one hand we can consider  $\der{E}:=\{\dder{E}_i\}_{i=0}^1$ as
		\[\xymatrix@R=20pt@C=20pt{1 \ar@{>->}[d]_{\delta^3_0}&& 1
	\ar@{>->}[d]_{\delta^3_0}\ar@{>->}[ll]_{\id{1}} \ar@{>->}[r]^{\delta^2_0} & 2
	\ar@{>->}@/^.25cm/[drrr]|(.32)\hole_(.4){i_2}
	\ar[dr]|(.33)\hole_{i_2} && 1 \ar@{>->}@/_.25cm/[dlll]^(.4){\delta^3_0}
	\ar@{>->}[dl]|(.33)\hole^{\delta^4_0}& 1 \ar@{>->}[d]^{\delta^4_0}\ar@{>->}[l]_{\id{1}}
	\ar@{>->}[rr]^{\id{1}} && 1 \ar@{>->}[d]^{\delta^4_0}\\3 && \ar@{>->}[ll]^{\id{3}}
	3 \ar@{>->}[rr]_{i_{3}}&& 4 && \ar@{>->}[ll]^{\id{4}} 4
	\ar@{>->}[rr]_{\id{4}}&& 4}
\]
and we have
\[
\delta^3_0=\id{3} \circ \delta^3_0
\qquad i_2=\id{4} \circ i_2
\qquad \delta^3_0= \id{3} \circ \delta^3_0
\qquad \delta^4_0= i_3\circ \delta^3_0\]

On the other hand, there is an isomorphism $t\colon 4\mto 4$ be the isomorphism which swap $1$ with $2$ and fix the other elements of $4$. Then we can also construct $\der{F}:=\{\dder{F}_i\}_{i=0}^1$ as below:
		\[\xymatrix@R=20pt@C=20pt{1 \ar@{>->}[d]_{\delta^3_0}&& 1
		\ar@{>->}[d]_{\delta^3_0}\ar@{>->}[ll]_{\id{1}} \ar@{>->}[r]^{\delta^2_0} & 2
		\ar@{>->}@/^.25cm/[drrr]|(.32)\hole_(.4){i_2}
		\ar[dr]|(.33)\hole_{i_2} && 1 \ar@{>->}@/_.25cm/[dlll]^(.4){\delta^3_0}
		\ar@{>->}[dl]|(.33)\hole^{\delta^4_0}& 1 \ar@{>->}[d]^{\delta^4_0}\ar@{>->}[l]_{\id{1}}
		\ar@{>->}[rr]^{\id{1}} && 1 \ar@{>->}[d]^{\delta^4_0}\\3 && \ar@{>->}[ll]^{\id{3}}
		3 \ar@{>->}[rr]_{i_{3}}&& 4 && \ar@{>->}[ll]^{\id{4}} 4
		\ar@{>->}[rr]_{t}&& 4}
	\]
	and also in this case we get a pre-switch since
	\[
	\delta^3_0=\id{3} \circ \delta^3_0
	\qquad i_2=t\circ i_2
	\qquad \delta^3_0= \id{3} \circ \delta^3_0
	\qquad \delta^4_0= i_3\circ \delta^3_0\]

	Now, by \Cref{thm:switch_uni} we get an abstraction equivalence fitting in the diagram below.
	\[\xymatrix@C=30pt@R=25pt{3 && \ar@{>->}[ll]_{\id{3}} 3 \ar@{>->}[rr]^{i_3}&& 4 && \ar@{>->}[ll]_{\id{4}} 4 \ar@{>->}[rr]^{\id{4}}&& 4\\1 \ar@{>->}[d]^{\delta^3_0} \ar@{>->}[u]_{\delta^3_0}&& 1 \ar@{>->}[d]^{\delta^3_0} \ar@{>->}[u]_{\delta^3_0}\ar@{>->}[ll]_{\id{1}} \ar@{>->}[r]^{\delta^2_0} & 2 \ar@{>->}@/^.35cm/[drrr]|(.315)\hole_(.4){i_2} \ar@{>->}[dr]|(.3)\hole_{i_2} \ar@{>->}@/_.35cm/[urrr]|(.315)\hole^(.4){i_2} \ar@{>->}[ur]|(.3)\hole^{i_2}& &1 \ar@{>->}@/_.35cm/[dlll]^(.4){\delta^3_0} \ar@{>->}[dl]|(.3)\hole^{\delta^4_0} \ar@{>->}@/^.35cm/[ulll]_(.4){\delta^3_0} \ar@{>->}[ul]|(.3)\hole_{\delta^4_0}& 1 \ar@{>->}[d]_{\delta^4_0} \ar@{>->}[u]^{\delta^4_0}\ar@{>->}[l]_{\id{1}} \ar@{>->}[rr]^{\id{1}} && 1 \ar@{>->}[d]_{\delta^4_0} \ar@{>->}[u]^{\delta^4_0}\\ \ar@/^.5cm/[uu]^{\id{3}} 3 && \ar[ll]^{\id{3}} 3 \ar@{>}@/^.5cm/[uu]^(.33){\phi_0}|\hole \ar[rr]_{i_3}&& 4 \ar@{>}[uu]_(.76){\psi_1}|(.45)\hole |(.55)\hole&& \ar@{>->}[ll]^{\id{4}} 4 \ar@{>}@/_.5cm/[uu]_(.33){\phi_1}|\hole\ar[rr]_{t}&& 4\ar@{>}@/_.5cm/[uu]_{\psi_2}}\]
	But then there is only one component of this abstraction equivalence different from the identity, namely the last one, which must coincide with $t$.
\end{example}

Another useful facts that we can establish is that being a pre-switch is invariant under abstraction equivalence, as made precise by the following lemma.

\begin{lemma}\label{lem:abs}
	Let $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be the following derivation:
	\[\xymatrix@R=20pt@C=20pt{L_0 \ar[d]_{m_0}&& K_0
		\ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
		\ar@{>}@/^.25cm/[drrr]|(.32)\hole_(.4){i_0}
		\ar[dr]|(.33)\hole_{h_0} && L_1 \ar@{>}@/_.25cm/[dlll]^(.4){i_1}
		\ar[dl]|(.33)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1}
		\ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0}
		D_0 \ar[rr]_{g_0}&& G_1 && \ar@{>->}[ll]^{f_1} D_1
		\ar[rr]_{g_1}&& G_2}
	\]
	and suppose that the derivation $\der{E}$ below is a pre-switch for $\der{D}$ along $(i_0,i_1)$.
	\[
	\xymatrix@R=20pt@C=20pt{
		{L_1} \ar[d]_{m_0'}
		&&  {K_1} \ar@{>->}[ll]_{l_1} \ar[r]^{r_1} \ar[d]_{k_0'}
		&  {R_1} \ar[dr]|(.33)\hole_{h_0'}  \ar@/^.25cm/@{>}_(.4){i_0'}|(.32)\hole[drrr]
		& & 
		{L_0}\ar[dl]|(.33)\hole^{m_1'} \ar@/_.25cm/@{>}^(.4){i_1'}[dlll] 
		&  {K_0} \ar@{>->}[l]_{l_0} \ar[rr]^{r_0} \ar[d]^{k_1'}
		& & {R_0} \ar[d]^{h_1'} \\		
		{G_0}
		& & {D_0'} \ar@{>->}[ll]^{f_0'} \ar[rr]_{g_0'}
		& &  {G'_1} 
		& &  {D_1'} \ar@{>->}[ll]^{f_1'} \ar[rr]_{g_1'}
		& & {G_2}  }
	\]

	Then the following hold true:
	\begin{enumerate}
		\item if $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ is an abstraction equivalence between $\der{D}$ and another derivation $\der{F}=\{\dder{F}_i\}_{i=0}^1$, then $\phi^{-1}_{G_0}*\der{E}*\phi_{G_2}$ is a pre-switch for $\der{F}$ along $(\phi_{D_1}\circ i_0, \phi_{D_0}\circ i_1 )$;
		\item if $\{\psi_X\}_{X\in \Deltamin(\der{E})}$ is an abstraction equivalence between $\der{E}$ and $\der{G}=\{\dder{G}_{i}\}_{i=0}^1$, then $\psi_{G_0}*\der{G}*\psi^{-1}_{G_2}$ is a pre-switch for $\der{D}$ along $(i_0, i_1)$.
	\end{enumerate}
	\begin{remark}Recall that, by \Cref{rem:seqabs}, $(\phi_{D_1}\circ i_0, \phi_{D_0}\circ i_1 )$ is an independence pair between $\dder{F}_0$ and $\dder{F}_1$.
	\end{remark}
\end{lemma}
\begin{proof}\begin{enumerate}
		\item To prove this, it is enough to compute:
		\[\begin{split}
			m_{\der{F},0}=&\phi_{G_0}\circ m_0\\=&\phi_{G_0}\circ f'_0\circ i'_1
		\end{split} \quad \begin{split}\phi_{G_0}\circ	m'_0=&\phi_{G_0}\circ f_0\circ i_1\\=&f_{\der{F},0}\circ \phi_{D_{0}}\circ i_1
	 \end{split} 
	\]
	\[\begin{split}
	h_{\der{F},1}=&\phi_{G_2}\circ h_1\\=&\phi_{G_2}\circ g_1' \circ i_0'
	\end{split} \qquad \begin{split}
		\phi_{G_2}\circ h'_1=&\phi_{G_2}\circ g_1\circ i_0\\=&g_{\der{F}, 1}\circ \phi_{D_1}\circ i_0
	\end{split}\]

		\item By \Cref{rem:seqabs} we know that $(\psi_{D'_1}\circ i'_0,\psi_{D'_0}\circ i'_1)$ is an independence pair between $\dder{G}_0$ and $\dder{G}_1$. By \Cref{rem:absequi} and again by \Cref{rem:seqabs} we conclude that  $(\psi_{D'_1}\circ i'_0,\psi_{D'_0}\circ i'_1)$ witnesses the sequential independence of the two steps making $\psi_{G_0}*\der{G}*\psi^{-1}_{G_2}$. Moreover:

>>>>

\chapter{Some examples of well-switching rewriting systems}

\subsubsection{Some examples of well-switching rewriting systems}
